Tropical Andes Taxonomic Diversity of Plants and Frugivores

### 2023-12-13 code/L2/L2_1_TropicalAndes_taxdiv.Rmd
add code to tranform sf to projected coordinate reference system, update plotting code to work with new crs, add more resolutions for mapping/plotting

# Polygons of countries
```{r}
worldMap <- ne_countries(scale = "medium", type = "countries", returnclass = "sf")
#polygon of Tropical Andes
TApoly <- worldMap %>% filter(sovereignt == "Bolivia" |sovereignt == "Ecuador" | sovereignt == "Venezuela" | sovereignt == "Colombia" | sovereignt == "Peru")
```

```{r}
TApoly <- st_transform(TApoly, crs = )
```

@@

```{r}
plants_sf_species <- plants.sf %>%
  group_by(species) %>%
  summarise()

@@

# Plot Base Map
```{r}
basePlot <-
ggplot() +
geom_sf(data = worldMap, fill = "white") +
geom_sf(data = TApoly) +
geom_sf(data = TropicalAndes_IUCNHabitat_Forest, fill = "forestgreen")+
labs(title = "Tropical Andes", x = "Latitude", y = "Longitude") +
coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
annotation_scale(location = "bl", width_hint = 0.5) +
annotation_north_arrow(location = "bl", which_north = "true", 
  pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
theme(panel.background = element_rect(fill = "lightblue"))

@@

# Plot points
```{r}
plantsPointsPlot <-
  ggplot() +
  geom_sf(data = worldMap, fill = "white") +
  geom_sf(data = TApoly) +
  geom_sf(data = plants_sf_species, pch = 21) +
  labs(title = "Tropical Andes Plant Occurrences", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
  theme(
    plot.background = element_rect(fill = "#f1f2f3"),
    panel.background = element_rect(fill = "lightblue"),
    panel.grid = element_blank(),
    line = element_blank(),
    rect = element_blank()
  )
plantsPointsPlot
```

```{r}
frugivoresPointsPlot <-
  ggplot() +
  geom_sf(data = worldMap, fill = "white") +
  geom_sf(data = TApoly) +
  geom_sf(data = frugivores_sf_species, pch = 21) +
  labs(title = "Tropical Andes Frugivore Occurrences", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
  theme(
    plot.background = element_rect(fill = "#f1f2f3"),
    panel.background = element_rect(fill = "lightblue"),
    panel.grid = element_blank(),
    line = element_blank(),
    rect = element_blank()
  )
frugivoresPointsPlot
```

# Make 0.5 Grid
```{r}
# make grid
TAGrid0.5 <- TApoly %>%
  st_make_grid(cellsize = c(0.5)) %>%
  st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
  st_cast("MULTIPOLYGON") %>%
  st_sf() %>%
  mutate(cellid = row_number())
```

# Plot Empty Grid
```{r}
# empty grid
gridPlot <-
  ggplot() +
  geom_sf(data = worldMap, fill = "white")+
  geom_sf(data = TropicalAndes_IUCNHabitat_Forest) +
  geom_sf(data = TAGrid0.5) +
  labs(title = "Tropical Andes Grid [0.5°]", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
  theme(
    plot.background = element_rect(fill = "#f1f2f3"),
    panel.background = element_rect(fill = "lightblue"),
    panel.grid = element_blank(),
    line = element_blank(),
    rect = element_blank()
  )
gridPlot
```

# Cell Richness
```{r}
plant_richness_grid0.5 <- TAGrid0.5 %>%
  st_join(plants_sf_species) %>%
  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
  group_by(cellid) %>%
  summarize(num_species = sum(overlap))
```

```{r}
frugivore_richness_grid0.5 <- TAGrid0.5 %>%
  st_join(frugivores_sf_species) %>%
  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
  group_by(cellid) %>%
  summarize(num_species = sum(overlap))
```

# Richness plots
```{r}
plantgridRichTA0.5 <-
  ggplot(plant_richness_grid0.5) +
  geom_sf(data = worldMap, fill = "white")+
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_sf(aes(fill = num_species), color = NA) +
  scale_fill_scico(palette = "romaO", direction = 1) +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
  theme(
    plot.background = element_rect(fill = "#f1f2f3"),
    panel.background = element_rect(fill = "lightblue"),
    panel.grid = element_blank(),
    line = element_blank(),
    rect = element_blank()
  ) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
  pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  labs(title = "Tropical Andes Plant Raw Richness", x = "Longitude", y = "Latitude", fill = "Number of species")
plantgridRichTA0.5
```

```{r}
frugivoregridRichTA0.5 <-
  ggplot(frugivore_richness_grid0.5) +
  geom_sf(data = worldMap, fill = "white")+
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_sf(aes(fill = num_species), color = NA) +
  scale_fill_scico(palette = "romaO", direction = 1) +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
  theme(
    plot.background = element_rect(fill = "#f1f2f3"),
    panel.background = element_rect(fill = "lightblue"),
    panel.grid = element_blank(),
    line = element_blank(),
    rect = element_blank()
  ) +   
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
  pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  labs(title = "Tropical Andes Frugivore Raw Richness", x = "Longitude", y = "Latitude", fill = "Number of species")
frugivoregridRichTA0.5
```

# Extract cell values
```{r}
plant_cellRichness0.5 <- plantgridRichTA0.5[["data"]]
frugivore_cellRichness0.5 <- frugivoregridRichTA0.5[["data"]]
```

```{r}
plot(plant_cellRichness0.5$num_species, frugivore_cellRichness0.5$num_species,
     main = "Plant Richness vs Frugivore Richness",
     xlab = "Plant Richness by cell", ylab = "Frugivore Richness by cell")
abline(lm(frugivore_cellRichness0.5$num_species ~ plant_cellRichness0.5$num_species))
summary(lm(frugivore_cellRichness0.5$num_species ~ plant_cellRichness0.5$num_species))
```

```{r}
hist(plant_cellRichness0.5$num_species,
     main = "Histogram of Plant Richness by cell",
     xlab = "Plant Richness by cell")
hist(frugivore_cellRichness0.5$num_species,
     main = "Histogram of Frugivore Richness by cell",
     xlab = "Frugivore Richness by cell")
```

```{r}
summary(plant_cellRichness0.5$num_species)
summary(frugivore_cellRichness0.5$num_species)
```

# Write data to csv
```{r}
write.csv(plant_cellRichness0.75, file.path(output_path,"TropicalAndes_GBIF_plant_richness_sub.csv"))
write.csv(frugivore_cellRichness0.75, file.path(output_path,"TropicalAndes_frugivore_richness_sub.csv"))
```

>>>>

# Polygons of countries
```{r}
worldMap <- ne_countries(scale = "medium", type = "countries", returnclass = "sf")
Americas <- ne_countries(continent = c("North America", "South America"), returnclass = "sf")
#polygon of Tropical Andes
TApoly <- worldMap %>% filter(sovereignt == "Bolivia" |sovereignt == "Ecuador" | sovereignt == "Venezuela" | sovereignt == "Colombia" | sovereignt == "Peru")
```

@@

# Transform to projected coordinate reference system (units from degrees to meters)
```{r}
Americas <- st_transform(Americas, 5389)
TApoly <- st_transform(TApoly, 5389)
TropicalAndes_IUCNHabitat_Forest <- st_transform(TropicalAndes_IUCNHabitat_Forest, 5389)
plants.sf <- st_transform(plants.sf, 5389)
frugivores.sf <- st_transform(frugivores.sf, 5389)
```

# Group by species
```{r}
plants_sf_species <- plants.sf %>%
  group_by(species) %>%
  summarise()

@@

# Plot Base Map
```{r}
basePlot <-
ggplot() +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly) +
  geom_sf(data = TropicalAndes_IUCNHabitat_Forest, fill = "forestgreen")+
  labs(title = "Tropical Andes", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

@@

# Plot points
```{r}
plantsPointsPlot <-
ggplot() +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly) +
  geom_sf(data = plants_sf_species, pch = 21) +
  labs(title = "Tropical Andes Plant Occurrences", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

plantsPointsPlot
```

```{r}
frugivoresPointsPlot <-
ggplot() +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly) +
  geom_sf(data = frugivores_sf_species, pch = 21) +
  labs(title = "Tropical Andes Frugivore Occurrences", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

frugivoresPointsPlot
```

# Plot TD/Richness

## 5 km resolution
### Make Grid
```{r}
#check units
st_crs(TApoly, parameters = TRUE)$units_gdal
st_crs(TApoly)
st_crs(TropicalAndes_IUCNHabitat_Forest)
```

```{r}
TAGrid_5 <- TApoly %>%
  st_make_grid(cellsize = c(5000)) %>%
  st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
  st_cast("MULTIPOLYGON") %>%
  st_sf() %>%
  mutate(cellid = row_number())
```

### Plot Empty Grid
```{r}
gridPlot <-
ggplot() +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TropicalAndes_IUCNHabitat_Forest) +
  geom_sf(data = TAGrid_5) +
  labs(title = "Tropical Andes Grid [5 km]", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

gridPlot
```

### Cell Richness
```{r}
plant_richness_grid_5 <- TAGrid_5 %>%
  st_join(plants_sf_species) %>%
  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
  group_by(cellid) %>%
  summarize(num_species = sum(overlap))
```

```{r}
frugivore_richness_grid_5 <- TAGrid_5 %>%
  st_join(frugivores_sf_species) %>%
  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
  group_by(cellid) %>%
  summarize(num_species = sum(overlap))
```

### Richness plots
```{r}
plantgridRichTA_5km <-
ggplot(plant_richness_grid_5) +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_sf(aes(fill = num_species), color = NA) +
  scale_fill_scico(palette = "romaO", direction = 1) +
  labs(title = "Tropical Andes Plant Richness [5 km]", x = "Longitude", y = "Latitude",
       fill = "Number of species") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

plantgridRichTA_5km
```

```{r}
frugivoregridRichTA_5km <-
ggplot(frugivore_richness_grid_5) +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_sf(aes(fill = num_species), color = NA) +
  scale_fill_scico(palette = "romaO", direction = 1) +
  labs(title = "Tropical Andes Frugivore Richness [5 km]", x = "Longitude", y = "Latitude",
       fill = "Number of species") +
    coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

frugivoregridRichTA_5km
```

## Extract cell values
```{r}
plant_cellRichness_5km <- plantgridRichTA_5km[["data"]]
frugivore_cellRichness_5km <- frugivoregridRichTA_5km[["data"]]
```

### Plant vs frugivore richness plot
```{r}
compare_richness_summary_5km <- summary(lm(frugivore_cellRichness_5km$num_species ~ plant_cellRichness_5km$num_species))
compare_richness_summary_5km

# save r-squared value
compare_richness_r2_5km <- compare_richness_summary_5km$r.squared
```

```{r}
plot(plant_cellRichness_5km$num_species, frugivore_cellRichness_5km$num_species,
     main = "Plant Richness vs Frugivore Richness [5 km]",
     xlab = "Plant Richness by cell", ylab = "Frugivore Richness by cell")
abline(lm(frugivore_cellRichness_5km$num_species ~ plant_cellRichness_5km$num_species))
```

```{r}
hist(plant_cellRichness_5km$num_species,
     main = "Histogram of Plant Richness by cell",
     xlab = "Plant Richness by cell")
hist(frugivore_cellRichness_5km$num_species,
     main = "Histogram of Frugivore Richness by cell",
     xlab = "Frugivore Richness by cell")
```

## 25 km 
### Make Grid
```{r}
TAGrid_25 <- TApoly %>%
  st_make_grid(cellsize = c(25000)) %>%
  st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
  st_cast("MULTIPOLYGON") %>%
  st_sf() %>%
  mutate(cellid = row_number())
```

### Plot Empty Grid
```{r}
gridPlot <-
ggplot() +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TropicalAndes_IUCNHabitat_Forest) +
  geom_sf(data = TAGrid_25) +
  labs(title = "Tropical Andes Grid [25 km]", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

gridPlot
```

### Cell Richness
```{r}
plant_richness_grid_25 <- TAGrid_25 %>%
  st_join(plants_sf_species) %>%
  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
  group_by(cellid) %>%
  summarize(num_species = sum(overlap))
```

```{r}
frugivore_richness_grid_25 <- TAGrid_25 %>%
  st_join(frugivores_sf_species) %>%
  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
  group_by(cellid) %>%
  summarize(num_species = sum(overlap))
```

### Richness plots
```{r}
plantgridRichTA_25km <-
ggplot(plant_richness_grid_25) +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_sf(aes(fill = num_species), color = NA) +
  scale_fill_scico(palette = "romaO", direction = 1) +
  labs(title = "Tropical Andes Plant Richness [25 km]", x = "Longitude", y = "Latitude",
       fill = "Number of species") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

plantgridRichTA_25km
```

```{r}
frugivoregridRichTA_25km <-
ggplot(frugivore_richness_grid_25) +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_sf(aes(fill = num_species), color = NA) +
  scale_fill_scico(palette = "romaO", direction = 1) +
  labs(title = "Tropical Andes Frugivore Richness [25 km]", x = "Longitude", y = "Latitude",
       fill = "Number of species") +
    coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

frugivoregridRichTA_25km
```

## Extract cell values
```{r}
plant_cellRichness_25km <- plantgridRichTA_25km[["data"]]
frugivore_cellRichness_25km <- frugivoregridRichTA_25km[["data"]]
```

### Plant vs frugivore richness plot
```{r}
compare_richness_summary_25km <- summary(lm(frugivore_cellRichness_25km$num_species ~ plant_cellRichness_25km$num_species))
compare_richness_summary_25km

# save r-squared value
compare_richness_r2_25km <- compare_richness_summary_25km$r.squared
```

```{r}
plot(plant_cellRichness_25km$num_species, frugivore_cellRichness_25km$num_species,
     main = "Plant Richness vs Frugivore Richness [25 km]",
     xlab = "Plant Richness by cell", ylab = "Frugivore Richness by cell")
abline(lm(frugivore_cellRichness_25km$num_species ~ plant_cellRichness_25km$num_species))
```

```{r}
hist(plant_cellRichness_25km$num_species,
     main = "Histogram of Plant Richness by cell",
     xlab = "Plant Richness by cell")
hist(frugivore_cellRichness_25km$num_species,
     main = "Histogram of Frugivore Richness by cell",
     xlab = "Frugivore Richness by cell")
```

## 50 km 
### Make Grid
```{r}
TAGrid_50 <- TApoly %>%
  st_make_grid(cellsize = c(50000)) %>%
  st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
  st_cast("MULTIPOLYGON") %>%
  st_sf() %>%
  mutate(cellid = row_number())
```

### Plot Empty Grid
```{r}
gridPlot <-
ggplot() +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TropicalAndes_IUCNHabitat_Forest) +
  geom_sf(data = TAGrid_50) +
  labs(title = "Tropical Andes Grid [50 km]", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

gridPlot
```

### Cell Richness
```{r}
plant_richness_grid_50 <- TAGrid_50 %>%
  st_join(plants_sf_species) %>%
  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
  group_by(cellid) %>%
  summarize(num_species = sum(overlap))
```

```{r}
frugivore_richness_grid_50 <- TAGrid_50 %>%
  st_join(frugivores_sf_species) %>%
  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
  group_by(cellid) %>%
  summarize(num_species = sum(overlap))
```

### Richness plots
```{r}
plantgridRichTA_50km <-
ggplot(plant_richness_grid_50) +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_sf(aes(fill = num_species), color = NA) +
  scale_fill_scico(palette = "romaO", direction = 1) +
  labs(title = "Tropical Andes Plant Richness [50 km]", x = "Longitude", y = "Latitude",
       fill = "Number of species") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

plantgridRichTA_50km
```

```{r}
frugivoregridRichTA_50km <-
ggplot(frugivore_richness_grid_50) +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_sf(aes(fill = num_species), color = NA) +
  scale_fill_scico(palette = "romaO", direction = 1) +
  labs(title = "Tropical Andes Frugivore Richness [50 km]", x = "Longitude", y = "Latitude",
       fill = "Number of species") +
    coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

frugivoregridRichTA_50km
```

## Extract cell values
```{r}
plant_cellRichness_50km <- plantgridRichTA_50km[["data"]]
frugivore_cellRichness_50km <- frugivoregridRichTA_50km[["data"]]
```

### Plant vs frugivore richness plot
```{r}
compare_richness_summary_50km <- summary(lm(frugivore_cellRichness_50km$num_species ~ plant_cellRichness_50km$num_species))
compare_richness_summary_50km

# save r-squared value
compare_richness_r2_50km <- compare_richness_summary_50km$r.squared
```

```{r}
plot(plant_cellRichness_50km$num_species, frugivore_cellRichness_50km$num_species,
     main = "Plant Richness vs Frugivore Richness [50 km]",
     xlab = "Plant Richness by cell", ylab = "Frugivore Richness by cell")
abline(lm(frugivore_cellRichness_50km$num_species ~ plant_cellRichness_50km$num_species))
```

```{r}
hist(plant_cellRichness_50km$num_species,
     main = "Histogram of Plant Richness by cell",
     xlab = "Plant Richness by cell")
hist(frugivore_cellRichness_50km$num_species,
     main = "Histogram of Frugivore Richness by cell",
     xlab = "Frugivore Richness by cell")
```

## 75 km 
### Make Grid
```{r}
TAGrid_75 <- TApoly %>%
  st_make_grid(cellsize = c(75000)) %>%
  st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
  st_cast("MULTIPOLYGON") %>%
  st_sf() %>%
  mutate(cellid = row_number())
```

### Plot Empty Grid
```{r}
gridPlot <-
ggplot() +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TropicalAndes_IUCNHabitat_Forest) +
  geom_sf(data = TAGrid_75) +
  labs(title = "Tropical Andes Grid [75 km]", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

gridPlot
```

### Cell Richness
```{r}
plant_richness_grid_75 <- TAGrid_75 %>%
  st_join(plants_sf_species) %>%
  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
  group_by(cellid) %>%
  summarize(num_species = sum(overlap))
```

```{r}
frugivore_richness_grid_75 <- TAGrid_75 %>%
  st_join(frugivores_sf_species) %>%
  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
  group_by(cellid) %>%
  summarize(num_species = sum(overlap))
```

### Richness plots
```{r}
plantgridRichTA_75km <-
ggplot(plant_richness_grid_75) +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_sf(aes(fill = num_species), color = NA) +
  scale_fill_scico(palette = "romaO", direction = 1) +
  labs(title = "Tropical Andes Plant Richness [75 km]", x = "Longitude", y = "Latitude",
       fill = "Number of species") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

plantgridRichTA_75km
```

```{r}
frugivoregridRichTA_75km <-
ggplot(frugivore_richness_grid_75) +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_sf(aes(fill = num_species), color = NA) +
  scale_fill_scico(palette = "romaO", direction = 1) +
  labs(title = "Tropical Andes Frugivore Richness [75 km]", x = "Longitude", y = "Latitude",
       fill = "Number of species") +
    coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

frugivoregridRichTA_75km
```

## Extract cell values
```{r}
plant_cellRichness_75km <- plantgridRichTA_75km[["data"]]
frugivore_cellRichness_75km <- frugivoregridRichTA_75km[["data"]]
```

### Plant vs frugivore richness plot
```{r}
compare_richness_summary_75km <- summary(lm(frugivore_cellRichness_75km$num_species ~ plant_cellRichness_75km$num_species))
compare_richness_summary_75km

# save r-squared value
compare_richness_r2_75km <- compare_richness_summary_75km$r.squared
```

# Compare r-squared by resolution
Plot
# Multi panel plots
Richness maps
# Write data to csv
```{r}
write.csv(plant_cellRichness0.75, file.path(output_path,"TropicalAndes_GBIF_plant_richness_sub.csv"))
write.csv(frugivore_cellRichness0.75, file.path(output_path,"TropicalAndes_frugivore_richness_sub.csv"))
```
